A quantitative framework for investigating the reliability of empirical network construction: supplemental information

### 2018-06-27 manuscript/CI_interactions.tex
spellchecked

    Representing an ecological community as a network summarizes both species composition and interactions between species. A tabulation of the nodes (species) and their relative abundances forms the basis for traditional metrics of community composition such as alpha diversity. A network framework combines this tabulation of nodes with a list of interactions (links between nodes) so that networks provide additional, higher-order information on community structure. Despite this additional information, empirical descriptions of ecological networks are usually limited to static representations of the communities they describe. That is, whether the network is assembled based on aggregated data, a single intensive ``snapshot" sample, or expert knowledge, interactions are assumed to occur deterministically wherever and whenever the community is observed~\citep{Olesen2011a}.

@@

      The choice of prior will depend on the information available. In general, it is best to use a prior taken from a similar system as emphasized above. If such data are not available, a prior based on many networks with a similar level of resolution could be useful, as could a prior based on a preliminary round of sampling (analogous to ``training data"). If there is no suitable informative prior, it is best to use an explicitly uninformative prior. While an uninformative prior will not greatly reduce variance about $\lambda$, it does clearly state our lack of prior knowledge and will not bias our conclusions.

@@

      The high spatiotemporal resolution and unusally high sampling effort of this dataset make it well suited for illustrating the difficulties in completely sampling a network. To show the gaps in sampling, we compared the frequencies of observed co-occurrences and interactions. We analysed both the \emph{Salix}-galler and galler-natural enemy components of the network but, for brevity, present only the latter here (see~\emph{Appendix S4} for \emph{Salix}-galler results).

@@

  Based on the framework described above, we can now offer some recommendations for improved descriptions of ecological interactions. First, sample sizes should be as large as possible in order to reduce detection uncertainty. Second, where sampling on the order of 30-50 observations per species pair is not possible (i.e., in most cases), researchers should acknowledge the varying levels of confidence surrounding interactions between different species pairs. Including the $n$ and $k$ values for each interaction will clearly indicate which unobserved interactions are most likely to be observed with further sampling and which estimates are more reliable. Third, the uncertainty around interactions should be incorporated in calculations of network properties. Re-sampling networks based on a probabilistic understanding of networks is straightforward and gives distributions for network properties rather than point estimates. This not only acknowledges the fact that interactions vary over time and space but will also facilitate comparisons between networks by adding confidence intervals to ezostimates of network properties. This will allow us to say whether networks have different structures \emph{and} whether those differences are greater than we would expect given the inherent variability of interactions. To facilitate the practical application of these recommendations, we provide all code used in this paper in the supplementary material.

@@

  \begin{figure}[h!]
    \caption{Three nested levels of uncertainty mean that an observed interaction matrix is unlikely to capture all of the interactions that can really occur at a site. Interaction uncertainty means that some inteactions that truly occur (black squares in True matrix) are less likely than others (grey squares; lighter grey is less likely) based on species traits. For example, lions are less likely to predate upon elephants than zebras. If our trait model is incomplete (e.g., if we neglect group hunting in lions), we might assume that elephants are too large to ever be prey for lions and assign this interaction a probability of 0.\\
    Process uncertainty describes the fact that interactions occur with different probabilities during any given sample. We show six example networks representing interactions observed over two days at three sites. Interactions which occur during fewer sampling days are less likely to be observed.\\
    Detection uncertainty reflects the fact that not all interactions are equally detectable. For example, interactions involving cryptic species are less likely to be seen.\\
    After applying these three layers of uncertainty, some interactions which truly occur are very unlikely to be observed. The observed matrix (lower right) is a subsample of the true matrix, with the probability of observing each interaction depending on the interaction, process, and detection uncertainty associated with it. Note that some low-probability interactions (e.g., lion predation on zebra) are included while others (e.g., lion predation on elephants) are not.}
    \label{conceptual_fig}
    \begin{center}
    \includegraphics*[height=.5\textheight]{figures/conceptual_fig.eps}
    \end{center}
    \end{figure}

@@

  \begin{figure}[h!]
    \caption{Imperfect detection of interactions increases the number of samples required to be confident that an interaction never occurs. Assume that two species cannot interact (i.e., $\lambda=0$) and that the number of observed interactions follows a binomial distribution. Here we show the upper bound (solid black line) of a 95\% Cloper-Pearson true credible interval for the interaction probability $\lambda$ for a pair of species that has been observed co-occurring $n$ times but never interacting ($k = 0$). The upper limit of the credible interval only reaches 0.1 (dashed, red line) with 35 observations. Thus, adding more observations is useful in controlling uncertainty, but the number of observations added must be very high. }
    \label{upper_limits}
    \begin{center}
    \includegraphics*[width=.8\textwidth]{figures/upper_limit_DG.eps}
    \end{center}
    \end{figure}

  \begin{figure}[h!]
    \caption{Using prior distributions based on a single site (Zillis in Graub\"{u}nden, Switzerland) from~\citet{Kopelke2017}, we calculate posterior distributions for the probability of interaction ($\lambda$) between two species that have not yet been observed interacting ($k = 0$). 
    Posterior distributions (curves) and 95\% credible intervals (lines at top of panell) for $\lambda$ narrow and approach zero as the number of observed co-occurrences ($n$) increases. Diamonds indicate the maximum likelihood estimator for the mean probability of interaction.
    Dashed lines indicate threshold probabilities of 0.01, 0.05, and 0.1. The number of samples required to obtain a 95\% credible interval below each threshold increases rapidly. It takes just over 20 observed co-occurrences to be 95\% confident that $\lambda<0.10$, approximately 50 co-occurrences to be 95\% confident that $\lambda<0.05$, and over 100 co-occurrences to be 95\% confident that $\lambda<0.01$.}
    \label{Salix_pdfs_cdfs}
    \begin{center}
    \includegraphics*[width=.8\textwidth]{figures/GP_pdfs_increasing_N_Zillis.eps}
    \end{center}
    \end{figure}

>>>>

    Representing an ecological community as a network summarises both species composition and interactions between species. A tabulation of the nodes (species) and their relative abundances forms the basis for traditional metrics of community composition such as alpha diversity. A network framework combines this tabulation of nodes with a list of interactions (links between nodes) so that networks provide additional, higher-order information on community structure. Despite this additional information, empirical descriptions of ecological networks are usually limited to static representations of the communities they describe. That is, whether the network is assembled based on aggregated data, a single intensive ``snapshot" sample, or expert knowledge, interactions are assumed to occur deterministically wherever and whenever the community is observed~\citep{Olesen2011a}.

@@

      The choice of prior will depend on the information available. In general, it is best to use a prior taken from a similar system as emphasised above. If such data are not available, a prior based on many networks with a similar level of resolution could be useful, as could a prior based on a preliminary round of sampling (analogous to ``training data"). If there is no suitable informative prior, it is best to use an explicitly uninformative prior. While an uninformative prior will not greatly reduce variance about $\lambda$, it does clearly state our lack of prior knowledge and will not bias our conclusions.

@@

      The high spatiotemporal resolution and unusually high sampling effort of this dataset make it well suited for illustrating the difficulties in completely sampling a network. To show the gaps in sampling, we compared the frequencies of observed co-occurrences and interactions. We analysed both the \emph{Salix}-galler and galler-natural enemy components of the network but, for brevity, present only the latter here (see~\emph{Appendix S4} for \emph{Salix}-galler results).

@@

  Based on the framework described above, we can now offer some recommendations for improved descriptions of ecological interactions. First, sample sizes should be as large as possible in order to reduce detection uncertainty. Second, where sampling on the order of 30-50 observations per species pair is not possible (i.e., in most cases), researchers should acknowledge the varying levels of confidence surrounding interactions between different species pairs. Including the $n$ and $k$ values for each interaction will clearly indicate which unobserved interactions are most likely to be observed with further sampling and which estimates are more reliable. Third, the uncertainty around interactions should be incorporated in calculations of network properties. Re-sampling networks based on a probabilistic understanding of networks is straightforward and gives distributions for network properties rather than point estimates. This not only acknowledges the fact that interactions vary over time and space but will also facilitate comparisons between networks by adding confidence intervals to estimates of network properties. This will allow us to say whether networks have different structures \emph{and} whether those differences are greater than we would expect given the inherent variability of interactions. To facilitate the practical application of these recommendations, we provide all code used in this paper in the supplementary material.

@@

  \begin{figure}[h!]
    \caption{Three nested levels of uncertainty mean that an observed interaction matrix is unlikely to capture all of the interactions that can really occur at a site. Interaction uncertainty means that some interactions that truly occur (black squares in True matrix) are less likely than others (grey squares; lighter grey is less likely) based on species traits. For example, lions are less likely to predate upon elephants than zebras. If our trait model is incomplete (e.g., if we neglect group hunting in lions), we might assume that elephants are too large to ever be prey for lions and assign this interaction a probability of 0.\\
    Process uncertainty describes the fact that interactions occur with different probabilities during any given sample. We show six example networks representing interactions observed over two days at three sites. Interactions which occur during fewer sampling days are less likely to be observed.\\
    Detection uncertainty reflects the fact that not all interactions are equally detectable. For example, interactions involving cryptic species are less likely to be seen.\\
    After applying these three layers of uncertainty, some interactions which truly occur are very unlikely to be observed. The observed matrix (lower right) is a subsample of the true matrix, with the probability of observing each interaction depending on the interaction, process, and detection uncertainty associated with it. Note that some low-probability interactions (e.g., lion predation on zebra) are included while others (e.g., lion predation on elephants) are not.}
    \label{conceptual_fig}
    \begin{center}
    \includegraphics*[height=.5\textheight]{figures/conceptual_fig.eps}
    \end{center}
    \end{figure}

@@

  \begin{figure}[h!]
    \caption{Imperfect detection of interactions increases the number of samples required to be confident that an interaction never occurs. Assume that two species cannot interact (i.e., $\lambda=0$) and that the number of observed interactions follows a binomial distribution. Here we show the upper bound (solid black line) of a 95\% Clopper-Pearson true credible interval for the interaction probability $\lambda$ for a pair of species that has been observed co-occurring $n$ times but never interacting ($k = 0$). The upper limit of the credible interval only reaches 0.1 (dashed, red line) with 35 observations. Thus, adding more observations is useful in controlling uncertainty, but the number of observations added must be very high. }
    \label{upper_limits}
    \begin{center}
    \includegraphics*[width=.8\textwidth]{figures/upper_limit_DG.eps}
    \end{center}
    \end{figure}

  \begin{figure}[h!]
    \caption{Using prior distributions based on a single site (Zillis in Graub\"{u}nden, Switzerland) from~\citet{Kopelke2017}, we calculate posterior distributions for the probability of interaction ($\lambda$) between two species that have not yet been observed interacting ($k = 0$). 
    Posterior distributions (curves) and 95\% credible intervals (lines at top of panel) for $\lambda$ narrow and approach zero as the number of observed co-occurrences ($n$) increases. Diamonds indicate the maximum likelihood estimator for the mean probability of interaction.
    Dashed lines indicate threshold probabilities of 0.01, 0.05, and 0.1. The number of samples required to obtain a 95\% credible interval below each threshold increases rapidly. It takes just over 20 observed co-occurrences to be 95\% confident that $\lambda<0.10$, approximately 50 co-occurrences to be 95\% confident that $\lambda<0.05$, and over 100 co-occurrences to be 95\% confident that $\lambda<0.01$.}
    \label{Salix_pdfs_cdfs}
    \begin{center}
    \includegraphics*[width=.8\textwidth]{figures/GP_pdfs_increasing_N_Zillis.eps}
    \end{center}
    \end{figure}

### 2018-06-27 manuscript/supplemental.tex
spellchecked

  To demonstrate the effects of nested interaction, process, and detection uncertainty on network structure, we can observe how the structure of randomly-generated networks changes as we add layers of uncertainty. We began by constructing 100 ``true" networks with 50 plant species, 50 galler species, and a connectance of 0.1. We ensured that each network had at least one link per plant and per galler. For each network, we calculated the mean number of links per plant, mean number of links per galler, and nestedness (NODF) as described in the main text. For simplicity, we assume that there is no interaction uncertainty in this system (i.e., all interactions are either 100\% feasible or 100\% infeasible based on species' traits).

  Taking each ``true" network as a starting point, we then applied process uncertainty. The probability $p$ of the occurrance of each interaction on a given sampling day was assigned a random probability drawn from the uniform distribution. As the probability that an interaction will occur during sampling also depends on the probability of the two species co-occurring, we also assigned a random number of sampling days on which each species co-occurred. The number of co-occurrances, $n$, was drawn from a uniform distribution and could be any integer between 0 and 30 (our simulated number of sampling days), inclusive. The number of occurrances, $k$, of each interaction was then drawn from a binomial distribution based on $n$ and $p$. If the number of occurrances during sampling was greater than 0 and the interaction occurred in the true network, we included the interaction in the ``process-filtered" network. As the inclusion of interactions in the process matrices is random, we created 100 process-filtered networks for each true network. For each network, we calculated the number of plants and gallers (after removing any species with no interactions), mean links per plant and per galler, connectance, and NODF.

@@

  With each additional layer of uncertainty, the number of plants and gallers included in the web decreased (state means here). The connectance and number of links per plant and per galler decreased significantly at each level (Fig.~\ref{sim_props}A,B,D). The effects of uncertainty on nestedness were less predictable. The process-filtered networks were significantly less nested than the original networks, but the variation in the detection-filtered webs was so great that there were no significant differences between the original and detection-filtered networks (Fig.~\ref{sim_props}C). This simulation demonstrates the fact that process and detection uncertainty mean that we are very unlikely to observe all of the interactions that can truly occur in a system, as well as the sometimes unpredictable effects this uncertainty can have on higher-order structures such as nestedness.

  Note that this simulation uses several simplifying assumptions. For instance, we treat interactions as independent and ignore the possibility for, for example, competition between gallers. A thorough exploration of the effects of interaction contingency, different distributions of co-occurrences, etc. would be an interesting topic for future work but is beyond the scope of the current study. We also neglect interaction uncertainty in this example. Interaction uncertainty could be included by, for example, modelling interactions between individuals of each species. In this framework, individuals could vary in some trait (e.g., body size) that affects the interactions in which they participate. Very small or very large individuals might not participate in some interactions that are feasbile for most members of the species.

  \begin{figure}
  \label{sim_props}
  \caption{We created 100 random networks containing 50 plants and 50 gallers, with a connectance of 0.1. For each of these ``original" networks, we created 100 process-filtered ("process") networks incorporating uncertainty around the occurrance of each interaction during sampling. For each ``process" network, we further created 100 detection-filtered ("detection") networks incorporating uncertainty around the detection of each interaction. For each network, we show the mean (+/- SD) connectance, links per plant, nestedness (NODF), and links per galler.}
  \begin{center}
  \includegraphics*[width=\textwidth]{figures/simulation_example.eps}
  \end{center}
  \end{figure}

>>>>

  To demonstrate the effects of nested interaction, process, and detection uncertainty on network structure, we can observe how the structure of randomly-generated networks changes as we add layers of uncertainty. We began by constructing 100 ``true" networks with 50 plant species, 50 galler species, and a connectance of 0.1. We ensured that each network had at least one link per plant and per galler. For each network, we calculated the mean number of links per plant, mean number of links per galler, and nestedness (NODF) as described in the main text. For simplicity, we assume that there is no interaction uncertainty in this system (i.e., all interactions are either 100\% feasible or 100\% unfeasible based on species' traits).

  Taking each ``true" network as a starting point, we then applied process uncertainty. The probability $p$ of the occurrence of each interaction on a given sampling day was assigned a random probability drawn from the uniform distribution. As the probability that an interaction will occur during sampling also depends on the probability of the two species co-occurring, we also assigned a random number of sampling days on which each species co-occurred. The number of co-occurrences, $n$, was drawn from a uniform distribution and could be any integer between 0 and 30 (our simulated number of sampling days), inclusive. The number of occurrences, $k$, of each interaction was then drawn from a binomial distribution based on $n$ and $p$. If the number of occurrences during sampling was greater than 0 and the interaction occurred in the true network, we included the interaction in the ``process-filtered" network. As the inclusion of interactions in the process matrices is random, we created 100 process-filtered networks for each true network. For each network, we calculated the number of plants and gallers (after removing any species with no interactions), mean links per plant and per galler, connectance, and NODF.

@@

  With each additional layer of uncertainty, the number of plants and gallers included in the web decreased ([[state means here]]). The connectance and number of links per plant and per galler decreased significantly at each level (Fig.~\ref{sim_props}A,B,D). The effects of uncertainty on nestedness were less predictable. The process-filtered networks were significantly less nested than the original networks, but the variation in the detection-filtered webs was so great that there were no significant differences between the original and detection-filtered networks (Fig.~\ref{sim_props}C). This simulation demonstrates the fact that process and detection uncertainty mean that we are very unlikely to observe all of the interactions that can truly occur in a system, as well as the sometimes unpredictable effects this uncertainty can have on higher-order structures such as nestedness.

  Note that this simulation uses several simplifying assumptions. For instance, we treat interactions as independent and ignore the possibility for, for example, competition between gallers. A thorough exploration of the effects of interaction contingency, different distributions of co-occurrences, etc. would be an interesting topic for future work but is beyond the scope of the current study. We also neglect interaction uncertainty in this example. Interaction uncertainty could be included by, for example, modelling interactions between individuals of each species. In this framework, individuals could vary in some trait (e.g., body size) that affects the interactions in which they participate. Very small or very large individuals might not participate in some interactions that are feasible for most members of the species.

  \begin{figure}
  \label{sim_props}
  \caption{We created 100 random networks containing 50 plants and 50 gallers, with a connectance of 0.1. For each of these ``original" networks, we created 100 process-filtered ("process") networks incorporating uncertainty around the occurrence of each interaction during sampling. For each ``process" network, we further created 100 detection-filtered ("detection") networks incorporating uncertainty around the detection of each interaction. For each network, we show the mean (+/- SD) connectance, links per plant, nestedness (NODF), and links per galler.}
  \begin{center}
  \includegraphics*[width=\textwidth]{figures/simulation_example.eps}
  \end{center}
  \end{figure}
